\textbf{SPECFEM3D Cartesian}\\ \textbf{User Manual}

### 2014-12-23 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
updates manual

You can add -DFORCE\_VECTORIZATION to the compiler options in \texttt{flags.guess} (for all compilers, except for IBM for which the syntax is -WF,-DFORCE\_VECTORIZATION )
to speed up the code in the fluid (acoustic) parts (only; FORCE\_VECTORIZATION support for elastic parts has been discontinued in the source code).
This works fine if (and only if) your computer always allocates a contiguous memory block for each allocatable array;
this is the case for most machines and most compilers, but not all. For more details see https://github.com/geodynamics/specfem3d/issues/81 .
To check if that option works fine on your machine, run the code with and without it for a model containing a significant fluid layer
(or entirely fluid) and make sure the seismograms are identical.\\

@@

\texttt{http://www.softeng.rl.ac.uk/blog/2010/jan/30/callgraph-fortran-doxygen/} ,

@@

\texttt{http://www.stack.nl/~dimitri/doxygen/manual/diagrams.html} ,

@@

\texttt{http://www.stack.nl/~dimitri/doxygen/manual/config.html} ,

@@

\texttt{http://www.stack.nl/~dimitri/doxygen/manual/index.html} .

@@

The basic steps in creating a load-balanced, partitioned mesh with
CUBIT are:
\begin{description}
\item [{1.}] setting up a hexahedral mesh with CUBIT,
\item [{2.}] exporting the CUBIT mesh into a SPECFEM3D Cartesian file format
and
\item [{3.}] partitioning the SPECFEM3D Cartesian mesh files for a chosen
number of cores.
\end{description}
Examples are provided in the SPECFEM3D Cartesian package in the subdirectory
\texttt{EXAMPLES/}. We strongly encourage you to contribute your own
example to this package by contacting the CIG Computational Seismology
Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.

@@

Prior to running and compiling this code, you have to export your
mesh in CUBIT to an ABAQUS (.inp) format. For example, export mesh
block IDs belonging to volumes in order to check the quality of the
hexahedral elements. You also have to determine a number of parameters
of your mesh, such as the number of nodes and number of elements and
modify the header of the \texttt{check\_mesh\_quality\_CUBIT\_Abaqus.f90}
source file. Then, in the main directory, type
\begin{lyxcode}
{\small make~xcheck\_mesh\_quality\_CUBIT\_Abaqus}{\small \par}
\end{lyxcode}
and use
\begin{lyxcode}
{\small ./bin/xcheck\_mesh\_quality\_CUBIT\_Abaqus}{\small \par}
\end{lyxcode}
to generate an %not supported: AVS output file (\texttt{\small AVS\_meshquality.inp} in AVS UCD format) or
OpenDX output file (\texttt{\small DX\_mesh\_quality.dx}{\small )
that can be used to investigate mesh quality, e.g. skewness of elements,
and a Gnuplot histogram (}\texttt{\small mesh\_quality\_histogram.txt}{\small )
that can be plotted with gnuplot (type `}\texttt{\small gnuplot plot\_mesh\_quality\_histogram.gnu}{\small ').
The histogram is also printed to the screen. Analyze that skewness
histogram of mesh elements to make sure no element has a skewness
above approximately 0.75, otherwise the mesh is of poor quality (and
if even a single element has a skewness value above 0.80, then you
must definitely improve the mesh). If you want to start designing
your own meshes, this tool is useful for viewing your creations. You
are striving for meshes with elements with `cube-like' dimensions,
e.g., the mesh should contain no very elongated or skewed elements.}{\small \par}

@@

The partitioning is done in serial for now (in the next release we
will provide a parallel version of that code). It needs to be run in the \texttt{bin/} directory because it expects the \texttt{../DATA/Par\_file}. The synopsis is:
\begin{lyxcode}
./xdecompose\_mesh~nparts~input\_directory~output\_directory
\end{lyxcode}
where
\begin{itemize}
\item \texttt{nparts} is the number of partitions, i.e., the number of cores
for the parallel simulations,
\item \texttt{input\_directory} is the directory which holds all the files
generated by the Python script \texttt{cubit2specfem3d.py} explained
in the previous Section~\ref{subsec:Exporting-the-Mesh}, e.g. \texttt{MESH/},
and
\item \texttt{output\_directory} is the directory for the output of this
partitioner which stores ACII-format files named like \texttt{proc{*}{*}{*}{*}{*}{*}\_Database}
for each partition. These files will be needed for creating the distributed
databases, and have to reside in the directory \texttt{LOCAL\_PATH}
specified in the main \texttt{Par\_file}, e.g. in directory \texttt{OUTPUT\_FILES/DATABASES\_MPI}.
Please see Chapter~\ref{cha:Creating-Distributed-Databases} for
further details.
\end{itemize}
Note that all the files generated by the Python script \texttt{cubit2specfem3d.py}
must be placed in the \texttt{input\_directory} folder before running
the program.

@@

In the main directory, type
\begin{lyxcode}
{\small make~xmeshfem3D}{\small \par}
\end{lyxcode}
If all paths and flags have been set correctly, the mesher should
now compile and produce the executable \texttt{bin/xmeshfem3D}. Please
note that \texttt{xmeshfem3D} must be called directly from the \texttt{bin/}
directory, as most of the binaries of the package.

@@

In the main directory, type
\begin{lyxcode}
{\small make~xgenerate\_databases}{\small \par}
\end{lyxcode}
Input for the program is provided through the main parameter file
\texttt{Par\_file}, which resides in the subdirectory \texttt{DATA}.
Please note that \texttt{xgenerate\_databases} must be called directly
from the \texttt{bin/} directory, as most of the binaries of the package.

@@

Now that you have successfully generated the databases, you are ready
to compile the solver. In the main directory, type
\begin{lyxcode}
{\small make~xspecfem3D}{\small \par}
\end{lyxcode}
Please note that \texttt{xspecfem3D} must be called directly from
the \texttt{bin/} directory, as most of the binaries of the package.

The solver needs three input files in the \texttt{DATA} directory
to run:
\begin{description}
\item [{\texttt{Par\_file}}] the main parameter file which was discussed
in detail in the previous Chapter~\ref{cha:Creating-Distributed-Databases},
\item [{\texttt{CMTSOLUTION}} or {\texttt{FORCESOLUTION}}] the earthquake
source parameter file or the force source parameter file, and
\item [{\texttt{STATIONS}}] the stations file.
\end{description}
Most parameters in the \texttt{Par\_file} should be set prior to running
the databases generation. Only the following parameters may be changed
after running \texttt{xgenerate\_databases}:
\begin{itemize}
\item the simulation type control parameters: \texttt{SIMULATION\_TYPE}
and \texttt{SAVE\_FORWARD}
\item the time step parameters \texttt{NSTEP} and \texttt{DT}
\item the absorbing boundary control parameter \texttt{PML\_CONDITIONS}
on condition that the\\
 \texttt{PML\_INSTEAD\_OF\_FREE\_SURFACE} flag remains unmodified
after running the databases generation.
\item the movie control parameters \texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME},
and \texttt{NTSTEPS\_BETWEEN\_FRAMES}
\item the ShakeMap\textregistered{}option \texttt{CREATE\_SHAKEMAP}
\item the output information parameters \texttt{MOVIE\_TYPE}, \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
and \texttt{NTSTEP\_BETWEEN\_OUTPUT\_}~\\
 \texttt{SEISMOS}
\item the \texttt{PRINT\_SOURCE\_TIME\_FUNCTION} flags
\end{itemize}
Any other change to the \texttt{Par\_file} implies rerunning both
the database generator \texttt{xgenerate\_databases} and the solver
\texttt{xspecfem3D}.

For any particular earthquake, the \texttt{CMTSOLUTION} file that
represents the point source may be obtained directly from the Harvard
Centroid-Moment Tensor (CMT) web page \urlwithparentheses{www.seismology.harvard.edu}.
It looks like this:
\begin{lyxcode}
\begin{figure}[H]
\noindent \begin{centering}
{\small \includegraphics[width=1\textwidth]{figures/Hollywood_CMT}
}
\par\end{centering}

\caption{\label{fig:CMTSOLUTION-file}\texttt{CMTSOLUTION} file based on the
format from the Harvard CMT catalog. \textbf{M} is the moment tensor,
$M_{0}${\small {} }is the seismic moment, and $M_{w}$ is the moment
magnitude.}
\end{figure}

\end{lyxcode}
The \texttt{CMTSOLUTION} file should be edited in the following way:
\begin{itemize}
\item Set the latitude or UTM $x$ coordinate, longitude or UTM $y$ coordinate, depth of the source (in km).
\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
will not run otherwise.) The time shift parameter would simply apply
an overall time shift to the synthetics, something that can be done
in the post-processing (see Section \ref{sec:Process-data-and-syn}).
\item For point-source simulations (see finite sources, page \pageref{To-simulate-a})
we recommend setting the source half-duration parameter \texttt{half
duration} equal to zero, which corresponds to simulating a step source-time
function, i.e., a moment-rate function that is a delta function. If
\texttt{half duration} is not set to zero, the code will use a Gaussian
(i.e., a signal with a shape similar to a `smoothed triangle', as
explained in \citet{KoTr02a} and shown in Fig~\ref{fig:gauss.vs.triangle})
source-time function with half-width \texttt{half duration}. We prefer
to run the solver with \texttt{half duration} set to zero and convolve
the resulting synthetic seismograms in post-processing after the run,
because this way it is easy to use a variety of source-time functions
(see Section \ref{sec:Process-data-and-syn}). \citet{KoTr02a} determined
that the noise generated in the simulation by using a step source
time function may be safely filtered out afterward based upon a convolution
with the desired source time function and/or low-pass filtering. Use
the serial code \texttt{convolve\_source\_timefunction.f90} and the
script \texttt{convolve\_source\_timefunction.csh} for this purpose,
or alternatively use signal-processing software packages such as SAC
\urlwithparentheses{www.llnl.gov/sac}. Type

@@

\end{itemize}
\begin{figure}
\noindent \begin{centering}
\includegraphics[width=2.5in]{figures/gauss_vs_triangle_mod}
\par\end{centering}

@@

We have also added the ability to run several calculations (several earthquakes)
in an embarrassingly-parallel fashion from within the same run;
this can be useful when using a very large supercomputer to compute
many earthquakes in a catalog, in which case it can be better from
a batch job submission point of view to start fewer and much larger jobs,
each of them computing several earthquakes in parallel.
To turn that option on, set parameter NUMBER\_OF\_SIMULTANEOUS\_RUNS
to a value greater than 1 in file setup/constants.h.in before
configuring and compiling the code.
When that option is on, of course the number of processor cores used to start
the code in the batch system must be a multiple of NUMBER\_OF\_SIMULTANEOUS\_RUNS,
all the individual runs must use the same number of processor cores,
which as usual is NPROC in the input file DATA/Par\_file,
and thus the total number of processor cores to request from the batch system
should be NUMBER\_OF\_SIMULTANEOUS\_RUNS $\times$ NPROC.
All the runs to perform must be placed in directories called run0001, run0002, run0003 and so on (with exactly four digits)
and you must create a link from the root directory of the code
to the first copy of the executable programs by typing " ln -s run0001/bin bin ".

@@

Faults need to be handled in a special way during mesh generation.
A fault surface must lie at the interface between elements (the mesh
must honor the fault surfaces). Moreover, a fault is made of two surfaces
in contact. Each of these two surfaces needs a separate set of nodes.
This approach is known as \textquotedbl{}split nodes\textquotedbl{}.
Currently faults can only be run with \texttt{xdecompose\_mesh }and\texttt{
CUBIT. xmeshfem3D }is not yet ready to handle faults.\\
To facilitate the mesh generation with split nodes in CUBIT, we need
to separate the two fault surfaces by a small distance, effectively
creating a tiny opening of the fault (Figure \ref{fig:examples.splitnodes},
\ref{fig:examples.splitnodes-surfacetrace}). Note that only the interior
of the fault must be opened, its edges must remain closed (except
the edge on the free surface). The fault is automatically closed later
by SPECFEM3D.\\

@@

The CUBIT scripts ({*}.jou and {*}.py) in the directory $\mathtt{EXAMPLES}$
generate more complicated meshes. The {*}.py files are Python scripts
that execute CUBIT commands and use the CUBIT-python interface for
SPECFEM3D (see next section). The Python language allows to define
and manipulate variables to parameterize the mesh. Alternatively,
the Python script can call a CUBIT journal file ({*}.jou), which looks
like the example above. Variables can be defined and manipulated there
using the $\mathtt{APREPRO}$ language built in CUBIT.\\

@@

\end{itemize}
Read the documents in the directory $\mathtt{EXAMPLES/*/description}$.
They contain a description of the example and additional instructions
to run it. Visualize the results with the Matlab scripts in the directory
$\mathtt{EXAMPLES/*/post}$

@@

Three additional input files are required in the \texttt{DATA} directory
for dynamic and kinematic faults. They are $\mathtt{Par\_file\_faults}$,
$\mathtt{FAULT\_STATIONS}$ and $\mathtt{input\_file.txt}$ (optional).
If the former does not exist, the code assumes that there are no faults.\\

@@

The purpose of the Kelvin-Voigt viscosity in the dynamic fault solver
is to damp spurious oscillations generated by the fault slip at frequencies
that are too high to be resolved by the mesh. The viscosity $\mathtt{eta}$
(in seconds) depends on the size of the elements on the fault. Here
is how to set it:
\begin{enumerate}
\item Determine the average linear size of the elements on the fault plane,
$\mathtt{h\_fault}$. Usually this value is prescribed by the user
during mesh generation. Otherwise it can be found by inspection of
the mesh inside the CUBIT GUI.
\item Use the Matlab function $\mathtt{utils/critical\_timestep.m}$ to
compute\\
 $\mathtt{dtc\_fault}=\mathtt{critical\_timestep\left(c_{p},h\_fault,ngll\right)}$.
\\
This is the critical time step in an elastic medium for a hypothetical
element of cubic shape with size equal to $\mathtt{h\_fault}$.
\item Set $\mathtt{eta}$ in $\mathtt{Par\_file\_faults}$ to (0.1 to 0.3)$\ensuremath{\times}\mathtt{dtc\_fault}$.
A larger $\mathtt{eta}$ damps high-frequencies more aggressively
but it might also affect lower frequencies and rupture speed.
\end{enumerate}
Viscosity reduces numerical stability: the critical timestep in a
simulation with Kelvin-Voigt damping needs to be smaller than that
in a purely elastic simulation. Here is how to set the time step accordingly:
\begin{enumerate}
\item Run a test simulation without viscosity ($\mathtt{eta}$=0 and only
a few time steps)
\item Look for the \textquotedbl{}maximum suggested time step\textquotedbl{}
in $\mathtt{OUTPUT\_FILES/output\_mesher.txt}$. This is the critical
timestep of a purely elastic simulation, $\mathtt{dtc\_bulk}$.
\item Reset the timestep of the simulation with a Kelvin-Voigt material
to a value smaller than\\
 $\mathtt{dtc\_kv}=\mathtt{eta}\left(\sqrt{1+\mathtt{dtc\_bulk^{2}}/\mathtt{eta^{2}}}-1\right)$
\end{enumerate}
Note that in general $\mathtt{dtc\_bulk}$ is smaller than $\mathtt{dtc\_fault}$,
because elements off the fault might be smaller or more distorted
than element faces on the fault.

@@

Several output files are saved in $\mathtt{OUTPUT\_FILES}$:\\

\begin{enumerate}
\item Seismograms for each station on the fault plane given in $\mathtt{DATA/FAUL\_STATIONS}$.
One output file is generated for each station, named after the station.
The files are ascii and start with a header (22 lines long) followed
by a data block with the following format, one line per time sample:\\

@@

\item Seismograms at stations in the bulk (out of the fault plane) given
in $\mathtt{DATA/STATIONS}$.
\item Rupture time files are named $\mathtt{Rupture\_time\_FAULT}$-$\mathtt{id}$.
One file is generated for each fault. The files are ascii and start
with a header (12 lines long) followed by a data block with the following
format, one line per fault node:

@@

The directories $\mathtt{EXAMPLES/*/post}$ contain additional Matlab
scripts to generate figures specific to each example.

@@

\begin{itemize}
\item Depending on what type of misfit function is used for the source inversion,
adjoint sources need to be computed from the original recorded seismograms
for the selected stations and saved in a sub-directory called \texttt{SEM/}
in the root directory of the code, with the format \texttt{STA.NT.BX?.adj}, where \texttt{STA},
\texttt{NT} are the station name and network code given in the \texttt{DATA/STATIONS\_ADJOINT}
file, and \texttt{BX?} represents the component name of a particular
adjoint seismogram. Please note that the band code can change depending
on your sampling rate (see Appendix~\ref{cha:channel-codes} for
further details).
\item The adjoint seismograms are in the same format as the original seismogram
(\texttt{STA.NT.BX?.sem?}), with the same start time, time interval
and record length.
\end{itemize}
\item Notice that even if you choose to time reverse only one component
from one specific station, you still need to supply all three components
because the code is expecting them (you can set the other two components
to be zero).
\item Also note that since time-reversal is done in the code itself, no
explicit time-reversing is needed for the preparation of the adjoint
sources, i.e., the adjoint sources are in the same forward time sense
as the original recorded seismograms.
\end{enumerate}
\item \textbf{Set the related parameters and run the adjoint simulation}\\
 In the \texttt{DATA/Par\_file}, set the two related parameters to
be \texttt{SIMULATION\_TYPE = 2} and \texttt{SAVE\_FORWARD = .false.}.
More conveniently, use the scripts \texttt{utils/change\_simulation\_type.pl}
to modify the \texttt{Par\_file} automatically (\texttt{change\_simulation\_type.pl
-a}). Then run the solver to launch the adjoint simulation.
\item \textbf{Collect the seismograms at the original source location}

>>>>

You can add \texttt{--enable-vectorization} to the configuration options to speed up the code in the fluid (acoustic) and elastic parts.
This works fine if (and only if) your computer always allocates a contiguous memory block for each allocatable array;
this is the case for most machines and most compilers, but not all. To disable this feature, use option \texttt{--disable-vectorization}. 
For more details see \href{https://github.com/geodynamics/specfem3d/issues/81}{github.com/geodynamics/specfem3d/issues/81} .
To check if that option works fine on your machine, run the code with and without it for an acoustic/elastic model and make sure the seismograms are identical.\\

@@

\url{http://www.softeng.rl.ac.uk/blog/2010/jan/30/callgraph-fortran-doxygen/},

@@

\url{http://www.stack.nl/~dimitri/doxygen/manual/diagrams.html},

@@

\url{http://www.stack.nl/~dimitri/doxygen/manual/config.html},

@@

\url{http://www.stack.nl/~dimitri/doxygen/manual/index.html}.

@@

\noindent
The basic steps in creating a load-balanced, partitioned mesh with
CUBIT are:
\begin{description}
\item [{1.}] setting up a hexahedral mesh with CUBIT,
\item [{2.}] exporting the CUBIT mesh into a SPECFEM3D Cartesian file format
and
\item [{3.}] partitioning the SPECFEM3D Cartesian mesh files for a chosen
number of cores.
\end{description}
Examples are provided in the SPECFEM3D Cartesian package in the subdirectory
\texttt{EXAMPLES/}. We strongly encourage you to contribute your own
example to this package by contacting the CIG Computational Seismology
Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.

@@

Prior to running and compiling this code, you have to export your
mesh in CUBIT to an ABAQUS (.inp) format. For example, export mesh
block IDs belonging to volumes in order to check the quality of the
hexahedral elements. You also have to determine a number of parameters
of your mesh, such as the number of nodes and number of elements and
modify the header of the \texttt{check\_mesh\_quality\_CUBIT\_Abaqus.f90}
source file in directory \texttt{src/check\_mesh\_quality\_CUBIT\_Abaqus/}.

\noindent
Then, in the main directory, type
\begin{lyxcode}
{\small make~xcheck\_mesh\_quality\_CUBIT\_Abaqus}{\small \par}
\end{lyxcode}
and use
\begin{lyxcode}
{\small ./bin/xcheck\_mesh\_quality\_CUBIT\_Abaqus}{\small \par}
\end{lyxcode}
to generate an %not supported: AVS output file (\texttt{\small AVS\_meshquality.inp} in AVS UCD format) or
OpenDX output file (\texttt{\small DX\_mesh\_quality.dx}{\small )
that can be used to investigate mesh quality, e.g. skewness of elements,
and a Gnuplot histogram (}\texttt{\small mesh\_quality\_histogram.txt}{\small )
that can be plotted with gnuplot (type `}\texttt{\small gnuplot plot\_mesh\_quality\_histogram.gnu}{\small ').
The histogram is also printed to the screen. Analyze that skewness
histogram of mesh elements to make sure no element has a skewness
above approximately 0.75, otherwise the mesh is of poor quality (and
if even a single element has a skewness value above 0.80, then you
must definitely improve the mesh). If you want to start designing
your own meshes, this tool is useful for viewing your creations. You
are striving for meshes with elements with `cube-like' dimensions,
e.g., the mesh should contain no very elongated or skewed elements.}{\small \par}

@@

The partitioning is done in serial for now (in the next release we
will provide a parallel version of that code). It needs to be run in the main directory because it expects the \texttt{./DATA/Par\_file}. The synopsis is:
\begin{lyxcode}
./bin/xdecompose\_mesh~nparts~input\_directory~output\_directory
\end{lyxcode}
where
\begin{itemize}
\item \texttt{nparts} is the number of partitions, i.e., the number of cores
for the parallel simulations,
\item \texttt{input\_directory} is the directory which holds all the files
generated by the Python script \texttt{cubit2specfem3d.py} explained
in the previous Section~\ref{subsec:Exporting-the-Mesh}, e.g. \texttt{./MESH/},
and
\item \texttt{output\_directory} is the directory for the output of this
partitioner which stores ACII-format files named like \texttt{proc{*}{*}{*}{*}{*}{*}\_Database}
for each partition. These files will be needed for creating the distributed
databases, and have to reside in the directory \texttt{LOCAL\_PATH}
specified in the main \texttt{Par\_file}, e.g. in directory \texttt{./OUTPUT\_FILES/DATABASES\_MPI}.
Please see Chapter~\ref{cha:Creating-Distributed-Databases} for
further details.
\end{itemize}
Note that all the files generated by the Python script \texttt{cubit2specfem3d.py}
must be placed in the \texttt{input\_directory} folder before running
the program.

@@

\noindent
In the main directory, type
\begin{lyxcode}
{\small make~xmeshfem3D}{\small \par}
\end{lyxcode}
If all paths and flags have been set correctly, the mesher should
now compile and produce the executable \texttt{bin/xmeshfem3D}. Please
note that \texttt{xmeshfem3D} must be called directly from the main
directory, as most of the binaries of the package.

@@

\noindent
In the main directory, type
\begin{lyxcode}
{\small make~xgenerate\_databases}{\small \par}
\end{lyxcode}
Input for the program is provided through the main parameter file
\texttt{Par\_file}, which resides in the subdirectory \texttt{DATA}.
Please note that \texttt{xgenerate\_databases} must be called directly
from the main directory, as most of the binaries of the package.

@@

Now that you have successfully generated the databases, you are ready
to compile the solver. In the main directory, type
\begin{lyxcode}
{\small make~xspecfem3D}{\small \par}
\end{lyxcode}
Please note that \texttt{xspecfem3D} must be called directly from
the main directory, as most of the binaries of the package.\\

\noindent
The solver needs three input files in the \texttt{DATA} directory
to run:
\begin{description}
\item [{\texttt{Par\_file}}] the main parameter file which was discussed
in detail in the previous Chapter~\ref{cha:Creating-Distributed-Databases},
\item [{\texttt{CMTSOLUTION}} or {\texttt{FORCESOLUTION}}] the earthquake
source parameter file or the force source parameter file, and
\item [{\texttt{STATIONS}}] the stations file.
\end{description}
Most parameters in the \texttt{Par\_file} should be set prior to running
the databases generation. Only the following parameters may be changed
after running \texttt{xgenerate\_databases}:
\begin{itemize}
\item the simulation type control parameters: \texttt{SIMULATION\_TYPE}
and \texttt{SAVE\_FORWARD}
\item the time step parameters \texttt{NSTEP} and \texttt{DT}
\item the absorbing boundary control parameter \texttt{PML\_CONDITIONS}
on condition that the\\
 \texttt{PML\_INSTEAD\_OF\_FREE\_SURFACE} flag remains unmodified
after running the databases generation.
\item the movie control parameters \texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME},
and \texttt{NTSTEPS\_BETWEEN\_FRAMES}
\item the ShakeMap\textregistered{}option \texttt{CREATE\_SHAKEMAP}
\item the output information parameters \texttt{MOVIE\_TYPE}, \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
and \texttt{NTSTEP\_BETWEEN\_OUTPUT\_}~\\
 \texttt{SEISMOS}
\item the \texttt{PRINT\_SOURCE\_TIME\_FUNCTION} flags
\end{itemize}
Any other change to the \texttt{Par\_file} implies rerunning both
the database generator \texttt{xgenerate\_databases} and the solver
\texttt{xspecfem3D}.

For any particular earthquake, the \texttt{CMTSOLUTION} file that
represents the point source may be obtained directly from the Harvard
Centroid-Moment Tensor (CMT) web page \urlwithparentheses{www.seismology.harvard.edu}.
It looks like the example shown in Fig. \ref{fig:CMTSOLUTION-file}.
%\begin{lyxcode}
\begin{figure}[htp]
\noindent 
\begin{centering}
{\small \includegraphics[width=0.8\textwidth]{figures/Hollywood_CMT}
}
\par
\end{centering}
\caption{\texttt{CMTSOLUTION} file based on the
format from the Harvard CMT catalog. \textbf{M} is the moment tensor,
$M_{0}${\small {} }is the seismic moment, and $M_{w}$ is the moment
magnitude.}
\label{fig:CMTSOLUTION-file}
\end{figure}
%\end{lyxcode}
The \texttt{CMTSOLUTION} file should be edited in the following way:
\begin{itemize}
\item Set the latitude or UTM $x$ coordinate, longitude or UTM $y$ coordinate, depth of the source (in km).
\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
will not run otherwise.) The time shift parameter would simply apply
an overall time shift to the synthetics, something that can be done
in the post-processing (see Section \ref{sec:Process-data-and-syn}).
\item For point-source simulations (see finite sources, page \pageref{To-simulate-a})
we recommend setting the source half-duration parameter \texttt{half
duration} equal to zero, which corresponds to simulating a step source-time
function, i.e., a moment-rate function that is a delta function. If
\texttt{half duration} is not set to zero, the code will use a Gaussian
(i.e., a signal with a shape similar to a `smoothed triangle', as
explained in \citet{KoTr02a} and shown in Fig~\ref{fig:gauss.vs.triangle})
source-time function with half-width \texttt{half duration}. We prefer
to run the solver with \texttt{half duration} set to zero and convolve
the resulting synthetic seismograms in post-processing after the run,
because this way it is easy to use a variety of source-time functions
(see Section \ref{sec:Process-data-and-syn}). \citet{KoTr02a} determined
that the noise generated in the simulation by using a step source
time function may be safely filtered out afterward based upon a convolution
with the desired source time function and/or low-pass filtering. Use
the serial code \texttt{convolve\_source\_timefunction.f90} and the
script \texttt{convolve\_source\_timefunction.csh} for this purpose,
or alternatively use signal-processing software packages such as SAC
\urlwithparentheses{www.llnl.gov/sac}. Type

@@

\end{itemize}
\begin{figure}[htp]
\noindent \begin{centering}
\includegraphics[width=3.5in]{figures/gauss_vs_triangle_mod}
\par\end{centering}

@@

We have also added the ability to run several calculations (several earthquakes)
in an embarrassingly-parallel fashion from within the same run;
this can be useful when using a very large supercomputer to compute
many earthquakes in a catalog, in which case it can be better from
a batch job submission point of view to start fewer and much larger jobs,
each of them computing several earthquakes in parallel.
To turn that option on, set parameter \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS}
to a value greater than 1 in file setup/constants.h.in before
configuring and compiling the code.
When that option is on, of course the number of processor cores used to start
the code in the batch system must be a multiple of \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS},
all the individual runs must use the same number of processor cores,
which as usual is \texttt{NPROC} in the input file \texttt{DATA/Par\_file},
and thus the total number of processor cores to request from the batch system
should be \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS $\times$ NPROC}.
All the runs to perform must be placed in directories called \texttt{run0001}, \texttt{run0002}, \texttt{run0003} and so on (with exactly four digits)
and you must create a link from the root directory of the code
to the first copy of the executable programs by typing \texttt{ln -s run0001/bin bin}.

@@

Faults need to be handled in a special way during mesh generation.
A fault surface must lie at the interface between elements (the mesh
must honor the fault surfaces). Moreover, a fault is made of two surfaces
in contact. Each of these two surfaces needs a separate set of nodes.
This approach is known as \textquotedbl{}split nodes\textquotedbl{}.
Currently faults can only be run with \texttt{xdecompose\_mesh} and
\texttt{CUBIT}. \texttt{xmeshfem3D} is not yet ready to handle faults.\\
To facilitate the mesh generation with split nodes in CUBIT, we need
to separate the two fault surfaces by a small distance, effectively
creating a tiny opening of the fault (Figure \ref{fig:examples.splitnodes},
\ref{fig:examples.splitnodes-surfacetrace}). Note that only the interior
of the fault must be opened, its edges must remain closed (except
the edge on the free surface). The fault is automatically closed later
by SPECFEM3D.\\

@@

The CUBIT scripts ({*}.jou and {*}.py) in the directory \texttt{EXAMPLES}
generate more complicated meshes. The {*}.py files are Python scripts
that execute CUBIT commands and use the CUBIT-python interface for
SPECFEM3D (see next section). The Python language allows to define
and manipulate variables to parameterize the mesh. Alternatively,
the Python script can call a CUBIT journal file ({*}.jou), which looks
like the example above. Variables can be defined and manipulated there
using the $\mathtt{APREPRO}$ language built in CUBIT.\\

@@

\end{itemize}
Read the documents in the directory \texttt{EXAMPLES/*/description}.
They contain a description of the example and additional instructions
to run it. Visualize the results with the Matlab scripts in the directory
$\mathtt{EXAMPLES/*/post}$

@@

Three additional input files are required in the \texttt{DATA/} directory
for dynamic and kinematic faults. They are $\mathtt{Par\_file\_faults}$,
$\mathtt{FAULT\_STATIONS}$ and $\mathtt{input\_file.txt}$ (optional).
If the former does not exist, the code assumes that there are no faults.\\

@@

The purpose of the Kelvin-Voigt viscosity in the dynamic fault solver
is to damp spurious oscillations generated by the fault slip at frequencies
that are too high to be resolved by the mesh. The viscosity $\mathtt{eta}$
(in seconds) depends on the size of the elements on the fault. Here
is how to set it:
\begin{enumerate}
\item Determine the average linear size of the elements on the fault plane,
$\mathtt{h\_fault}$. Usually this value is prescribed by the user
during mesh generation. Otherwise it can be found by inspection of
the mesh inside the CUBIT GUI.
\item Use the Matlab function $\mathtt{utils/critical\_timestep.m}$ to
compute\\
 $\mathtt{dtc\_fault}=\mathtt{critical\_timestep\left(c_{p},h\_fault,ngll\right)}$.
\\
This is the critical time step in an elastic medium for a hypothetical
element of cubic shape with size equal to $\mathtt{h\_fault}$.
\item Set $\mathtt{eta}$ in $\mathtt{Par\_file\_faults}$ to (0.1 to 0.3)$\ensuremath{\times}\mathtt{dtc\_fault}$.
A larger $\mathtt{eta}$ damps high-frequencies more aggressively
but it might also affect lower frequencies and rupture speed.
\end{enumerate}
Viscosity reduces numerical stability: the critical timestep in a
simulation with Kelvin-Voigt damping needs to be smaller than that
in a purely elastic simulation. Here is how to set the time step accordingly:
\begin{enumerate}
\item Run a test simulation without viscosity ($\mathtt{eta}$=0 and only
a few time steps)
\item Look for the \textquotedbl{}maximum suggested time step\textquotedbl{}
in \texttt{OUTPUT\_FILES/output\_mesher.txt}. This is the critical
timestep of a purely elastic simulation, $\mathtt{dtc\_bulk}$.
\item Reset the timestep of the simulation with a Kelvin-Voigt material
to a value smaller than\\
 $\mathtt{dtc\_kv}=\mathtt{eta}\left(\sqrt{1+\mathtt{dtc\_bulk^{2}}/\mathtt{eta^{2}}}-1\right)$
\end{enumerate}
Note that in general $\mathtt{dtc\_bulk}$ is smaller than $\mathtt{dtc\_fault}$,
because elements off the fault might be smaller or more distorted
than element faces on the fault.

@@

Several output files are saved in \texttt{OUTPUT\_FILES/}:\\

\begin{enumerate}
\item Seismograms for each station on the fault plane given in \texttt{DATA/FAUL\_STATIONS}.
One output file is generated for each station, named after the station.
The files are ascii and start with a header (22 lines long) followed
by a data block with the following format, one line per time sample:\\

@@

\item Seismograms at stations in the bulk (out of the fault plane) given
in \texttt{DATA/STATIONS}.
\item Rupture time files are named $\mathtt{Rupture\_time\_FAULT}$-$\mathtt{id}$.
One file is generated for each fault. The files are ascii and start
with a header (12 lines long) followed by a data block with the following
format, one line per fault node:

@@

The directories \texttt{EXAMPLES/*/post} contain additional Matlab
scripts to generate figures specific to each example.

@@

\begin{itemize}
\item Depending on what type of misfit function is used for the source inversion,
adjoint sources need to be computed from the original recorded seismograms
for the selected stations and saved in a sub-directory called \texttt{SEM/}
in the root directory of the code, with the format \texttt{NT.STA.BX?.adj}, where \texttt{NT},
\texttt{STA} are the network code and station name given in the \texttt{DATA/STATIONS\_ADJOINT}
file, and \texttt{BX?} represents the component name of a particular
adjoint seismogram. Please note that the band code can change depending
on your sampling rate (see Appendix~\ref{cha:channel-codes} for
further details).
\item The adjoint seismograms are in the same format as the original seismogram
(\texttt{NT.STA.BX?.sem?}), with the same start time, time interval
and record length.
\end{itemize}
\item Notice that even if you choose to time reverse only one component
from one specific station, you still need to supply all three components
because the code is expecting them (you can set the other two components
to be zero).
\item Also note that since time-reversal is done in the code itself, no
explicit time-reversing is needed for the preparation of the adjoint
sources, i.e., the adjoint sources are in the same forward time sense
as the original recorded seismograms.
\end{enumerate}
\item \textbf{Set the related parameters and run the adjoint simulation}\\
 In the \texttt{DATA/Par\_file}, set the two related parameters to
be \texttt{SIMULATION\_TYPE = 2} and \texttt{SAVE\_FORWARD = .false.}.
More conveniently, use the scripts \texttt{utils/change\_simulation\_type.pl}
to modify the \texttt{Par\_file} automatically (\texttt{change\_simulation\_type.pl
-a}). Then run the solver to launch the adjoint simulation.
\item \textbf{Collect the seismograms at the original source location}
